### 2024-06-17 sections/2-methods.tex
shrunk to 700 lines

We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
A distance value of zero defines an identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among the points.
We also require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
In addition to these constraints, if the distance function obeys the triangle inequality (i.e. $f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while cosine distance is not a metric because it violates the triangle inequality (e.g., consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).

@@

CAKES derives its advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
While we sometimes use Euclidean notions, such as voids, volumes and proximity to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.

CAKES exploits the low LFD of such datasets to accelerate search.
We define LFD at some length scale around a point in the dataset as:

\begin{equation}
    \frac{\text{log} \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right) }{\text{log} \left( \frac{r_1}{r_2} \right) }
    \label{eq:methods:lfd-original}
\end{equation}
where $B_X(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
Intuitively, LFD measures the rate of change in the number of points in a ball of radius $r$ around a point $q$ as $r$ increases. When the vast majority of points in the dataset have low ($\ll D$) LFD, we can overload terminology to say that the dataset has low LFD.

We stress that this concept differs from the \textit{embedding dimension} of a dataset.
To illustrate the difference, consider the SILVA 18S rRNA dataset which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
However, due to physical constraints (namely, biological evolution and biochemistry), the data are constrained to a lower-dimensional manifold within this embedding space.
LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.

@@

CLAM builds a binary tree of clusters using a divisive hierarchical clustering algorithm.

For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
Using these distances, we compute the \textit{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
This geometric median is the center of $C$.

The \textit{radius} of $C$ is the maximum distance from the \textit{center} to any other point in $C$.
The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from the left pole is designated the \textit{right pole}.

We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster which are closer to the left pole than to the right pole, and the right child contains all points in the cluster which are closer to the right pole than to the left pole.
Without loss of generality, we assign to the left child those points which are equidistant from the two poles.

Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria, for example, minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc.
This process is described in Algorithm \ref{alg:methods:partition}.

During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-simplified}.

@@

After building the tree, we reorder the dataset so that the points are stored in the order that they would be visited during a depth-first traversal of the tree.

In CHESS, each cluster stored a list of indices into the dataset.
This list was used to retrieve the clusters' points during search.
Although this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high.
With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree.
Assuming a balanced tree, and thus $\mathcal{O}(\log n)$ depth, this approach had a memory overhead of $\mathcal{O}(n \log n)$.

One may improve this memory overhead to $\mathcal{O}(n)$ by only storing indices at the leaf clusters.
This approach, however, introduces a time cost of $\mathcal{O}(n \log n)$ whenever we need to find the indices for a non-leaf cluster (because it requires a traversal of the subtree rooted at that cluster).

In this work, we introduce a new approach wherein, after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset.
The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.
A left child has the same offset as that of its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.

With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers the same time complexity as CHESS with $\mathcal{O}(n)$ space.

@@

We define a \textit{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains only duplicates of the same point (i.e., has cardinality greater than 1 but contains only one \textit{unique} point).
A singleton clearly has zero radius, and so $\delta = \delta^{-} = \delta^{+}$.
Hence, we overload the above notation to also refer to the distance from a query to an individual point.

@@

To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.

Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.

@@

\begin{itemize}
    \item While the top priority element $Q$ is not a leaf, remove it from $Q$ and add its children to $Q$.
    \item Remove the top priority element (a leaf) from $Q$ and add all its points to $H$.
    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
\end{itemize}

@@

Due to their similarity, we combine the complexity analyses of the two Sieve methods into one section.
For these methods we, once again, use the terminology of tree-search and leaf-search.
Tree-search navigates the cluster tree and adds clusters to $Q$.
Leaf-search exhaustively searches some of the clusters in $Q$ to find the $k$ nearest neighbors.

@@

The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is

@@

For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the complexity of the priority queue operations.
Thus, the asymptotic complexity is

@@

To elaborate on the augmentation process, we start with an original dataset from the ANN-benchmarks suite.
Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level, and $m$ be a user-specified integer multiplier.
For each datum $x \in X$, we create $m - 1$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the Euclidean distance from $x$ to the origin.
We construct a random vector $r$ of $d$ dimensions in the hyper-sphere of radius $\epsilon$ centered at the origin.
We then add $r$ to $x$ to get a new point $x'$.
Since $||r|| \leq \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e., $x'$ is within a distance $\epsilon$ of $x$).
This produces a new dataset $X'$ with $|X'| = m \cdot |X|$.

This augmentation process preserves the topological structure of the original dataset, but increases its cardinality by a factor of $m$, allowing us to isolate the effect of cardinality on search performance from that of other factors such as dimensionality, choice of metric, or the topological structure of the dataset.

>>>>

We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
A distance of zero defines identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among points.
We also require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
In addition to these constraints, if the distance function obeys the triangle inequality (i.e. $f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while cosine distance is not a metric because it violates the triangle inequality (e.g., consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).

@@

CAKES derives its advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
While we sometimes use Euclidean notions to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.
CAKES exploits the low LFD of such datasets to accelerate search.
We define LFD at some length scale around a point in the dataset as:

\begin{equation}
    \frac{\text{log} \left( \frac{|B_X(q, r_1)|}{|B_X(q, r_2)|} \right) }{\text{log} \left( \frac{r_1}{r_2} \right) }
    \label{eq:methods:lfd-original}
\end{equation}
where $B_X(q, r)$ is the set of points contained in the metric ball of radius $r$ centered at a point $q$ in the dataset $\textbf{X}$.
Intuitively, LFD measures the rate of change in the number of points in a ball of radius $r$ around a point $q$ as $r$ increases. When the vast majority of points in the dataset have low ($\ll D$) LFD, we can simply say that the dataset has low LFD.
We stress that this concept differs from the \textit{embedding dimension} of a dataset.
To illustrate the difference, consider the SILVA 18S rRNA dataset which contains genomic sequences with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
However, physical constraints (namely, biological evolution and biochemistry) constrain the data to a lower-dimensional manifold within this space.
LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.

@@

CLAM builds a binary tree of clusters using a divisive hierarchical clustering algorithm.
For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
Using these distances, we compute the \textit{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
This geometric median is the center of $C$.

The \textit{radius} of $C$ is the maximum distance from the \textit{center} to any other point in $C$.
The point which is responsible for that radius (i.e., the furthest point from the center) is designated the \textit{left pole} and the point which is furthest from the left pole is designated the \textit{right pole}.
We then partition the cluster into a \textit{left child} and a \textit{right child}, where the left child contains all points in the cluster which are closer to the left pole than to the right pole, and the right child contains all points in the cluster which are closer to the right pole than to the left pole.
Without loss of generality, we assign to the left child those points which are equidistant from the two poles.
Starting from a root cluster containing the entire dataset, we repeat this procedure until each leaf contains only one datum, or we meet some other user-specified stopping criteria, for example, minimum cluster radius, minimum cluster cardinality, maximum tree depth, etc.
This process is described in Algorithm \ref{alg:methods:partition}.
During the partitioning process, we also compute (and cache) the LFD of each cluster using Equation~\ref{eq:methods:lfd-simplified}.

@@

%After building the tree, we reorder the dataset so that the points are stored in the order that they would be visited during a depth-first traversal of the tree.

In CHESS, each cluster stored a list of indices into the dataset.
This list was used to retrieve the clusters' points during search.
Although this approach allowed us to retrieve the points in constant time, its memory cost was prohibitively high.
With a dataset of cardinality $n$ and each cluster storing a list of indices for its points, we stored a total of $n$ indices at each depth in the tree.
Assuming a balanced tree, and thus $\mathcal{O}(\log n)$ depth, this approach had a memory overhead of $\mathcal{O}(n \log n)$.
In this work, we introduce a new approach wherein, after building the cluster tree, we reorder the dataset so that points are stored in a depth-first order.
Then, within each cluster, we need only store its \textit{cardinality} and an \textit{offset} to access its points from the dataset, based on a useful invariant.
The root cluster has an offset of zero and a cardinality equal to the number of points in the dataset.
A left child has the same offset its parent, and the corresponding right child has an offset equal to the left child's offset plus the left child's cardinality.
With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers the same time complexity as CHESS with $\mathcal{O}(n)$ space.

@@

We define a \textit{singleton} as a cluster which either contains a single point (i.e., has cardinality 1) or which contains only duplicates of the same point.
A singleton clearly has zero radius, and so $\delta = \delta^{-} = \delta^{+}$.
Hence, we overload the above notation to also refer to the distance from a query to an individual point.

@@

To check whether $d \leq \rho$, we note that $d = \overline{m q' \vphantom{l}} = \overline{m r \vphantom{l}} - \overline{q' r \vphantom{l}} = \frac{\overline{l r}}{2} - \overline{q' r \vphantom{l}}$.
Let $\theta$ denote $\angle l r q$, as shown in Figure~\ref{fig:methods:overlapping-children}.
By the Law of Cosines on $\triangle l r q$, we have that $\text{cos}(\theta) = \tfrac{\overline{l r}^2 + \ \overline{q r \vphantom{l}}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r} \cdot \overline{q r \vphantom{l}}}$.
Since $\triangle r q q'$ is a right triangle, we also have that $\text{cos}(\theta) = \tfrac{\overline{q' r \vphantom{l}}}{\overline{q r \vphantom{l}}}$.
Combining the previous two equations and solving for $\overline{q' r \vphantom{l}}$, we have that $\overline{q' r \vphantom{l}} = \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}}$.
Substituting for $\overline{q' r \vphantom{l}}$ in the equation for $d$, we have that $d = \tfrac{\overline{l r}}{2} - \tfrac{\overline{q r \vphantom{l}}^2 + \ \overline{l r}^2 - \ \overline{q l}^2}{2 \cdot \overline{l r}} = \tfrac{\overline{q l}^2 - \overline{q r \vphantom{l}}^2}{2 \cdot \overline{l r}}$.

@@

\begin{itemize}
    \item While the top-priority element is not a leaf, remove it from $Q$ and add its children to $Q$.
    \item Remove the top-priority element (a leaf) from $Q$ and add all its points to $H$.
    \item If $H$ has more than $k$ points, remove points from $H$ until $|H| = k$.
\end{itemize}

@@

Due to their similarity, we combine the complexity analyses of both Sieve methods.
For these methods we again use the terminology of tree-search and leaf-search.
Tree-search navigates the cluster tree and adds clusters to $Q$.
Leaf-search exhaustively searches some of the clusters in $Q$ to find the $k$ nearest neighbors.

@@

The asymptotic complexity of Breadth-First Sieve is dominated by the QuickSelect algorithm to calculate $\tau$.
Since this method is log-linear in the length of $Q$, and $Q$ contains the clusters from tree-search and the points from leaf-search, we see that the asymptotic complexity is:

@@

For Depth-First Sieve, since we use two priority queues, the asymptotic complexity is dominated by the priority queue operations.
Thus, the complexity is:

@@

To elaborate on the augmentation process, we start with an original dataset from the ANN-benchmarks suite.
Let $X$ be the dataset, $d$ be its dimensionality, $\epsilon$ be a user-specified noise level, and $m$ be a user-specified integer multiplier.
For each datum $x \in X$, we create $m - 1$ new data points within a distance $\epsilon \cdot ||x||$ of $x$ where $||x||$ is the Euclidean distance from $x$ to the origin.
We construct a random vector $r$ of $d$ dimensions in the hyper-sphere of radius $\epsilon$ centered at the origin.
We then add $r$ to $x$ to get a new point $x'$.
Since $||r|| \leq \epsilon$, we have that $||x - x'|| \leq \epsilon$ (i.e., $x'$ is within a distance $\epsilon$ of $x$).
This produces a new dataset $X'$ with $|X'| = m \cdot |X|$.
This augmentation process preserves the topological structure of the original dataset, but increases its cardinality by a factor of $m$, allowing us to isolate the effect of cardinality on search performance from that of other factors such as dimensionality, choice of metric, or the topological structure of the dataset.

### 2024-06-17 sections/3-datasets.tex
shrunk to 700 lines

We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}, using the associated distance function.
Table~\ref{tab:datasets:summary} summarizes the properties of these datasets.

The benchmarks for Fashion-Mnist, Glove-25, Sift, and Random (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}.
The OS kernel was Ubuntu 22.04.3-Ubuntu SMP.
The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.

The benchmarks for SILVA and RadioML were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
The OS kernel was Manjaro Linux 5.15.130-1-MANJARO.
The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.

@@

In addition to benchmarks on datasets from the ANN-Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{sec:methods:synthetic-data}. 
In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Multiplier'' in Table~\ref{tab:results:qps-and-recall}) increases.

We also benchmarked on purely random (i.e., not an augmented version of a known dataset), datasets of various cardinalities.
For this, we used a base cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset; hereafter, we refer to the random dataset with cardinality 1,000,000 and dimensionality 128 as ``Random.'' 
This benchmark allows us to isolate the effect of a manifold structure (which we expect to be absent in a purely random dataset) on the performance of the CAKES' algorithms.

@@

To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}.
This dataset contains ribosomal RNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
Since this dataset contains many redundant, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition criterion in Algorithm~\ref{alg:methods:partition}.

@@

To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset~\cite{oshea2018radioml}.
The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
Specifically, it comprises 24 modulation modes at 26 different signal-to-noise ratio (SNR) levels ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time-series of complex-valued numbers).
We used a subset of this dataset, containing 97,920 samples at 10dB SNR and used the other 8,576 samples at 10dB SNR as a hold-out set of queries.
We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.

>>>>

We benchmark on a variety of datasets from the ANN-benchmarks suite~\cite{Aumller2018ANNBenchmarksAB}.
Table~\ref{tab:datasets:summary} summarizes these datasets.
The benchmarks for Fashion-Mnist, Glove-25, Sift, and Random (and for their synthetic augmentations) were conducted on an Amazon AWS (EC2) \texttt{r6i.16xlarge} instance, with a 64-core Intel Xeon Platinum 8375C CPU 2.90GHz processor, 512GB RAM.
This is the same configuration used in~\cite{Aumller2018ANNBenchmarksAB}.
The OS kernel was Ubuntu 22.04.3-Ubuntu SMP.
The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.
The benchmarks for SILVA and RadioML were conducted on an Intel Xeon E5-2690 v4 CPU @ 2.60GHz with 512GB RAM.
The OS kernel was Manjaro Linux 5.15.130-1-MANJARO.
The Rust compiler was Rust 1.72.0, and the Python interpreter version was 3.9.16.

@@

In addition to benchmarks on datasets from the ANN-Benchmarks suite, we also benchmarked on synthetic augmentations of these real datasets, using the process described in \ref{sec:methods:synthetic-data}.
In particular, we use a noise tolerance $\epsilon = 0.01$ and explore the scaling behavior as the cardinality multiplier (referred to as ``Multiplier'' in Table~\ref{tab:results:qps-and-recall}) increases.
We also benchmarked on purely randomly-generated datasets of various cardinalities.
For this, we used a base cardinality of 1,000,000 and a dimensionality of 128 to match the Sift dataset; hereafter, we refer to the random dataset with cardinality 1,000,000 and dimensionality 128 as ``Random.''
This benchmark allows us to isolate the effect of a manifold structure (which we expect to be absent in a purely random dataset) on the performance of the CAKES' algorithms.

@@

To demonstrate CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}.
This dataset contains ribosomal RNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
Since this dataset contains many redundant, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition criterion in Algorithm~\ref{alg:methods:partition}.

@@

As another example of an exotic distance function, we benchmarked CAKES on the RadioML dataset~\cite{oshea2018radioml}.
The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
Specifically, it comprises 24 modulation modes at 26 different signal-to-noise ratio (SNR) levels ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time-series of complex-valued numbers).
We used a subset of this dataset, containing 97,920 samples at 10dB SNR and used the other 8,576 samples at 10dB SNR as a hold-out set of queries.
We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.
